A community-maintained standard library of population genetic models

### 2019-12-20 popSimManu.tex

\subsection*{The \stdpopsim Catalog}
The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing, the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}, 
and we have models in development for \emph{Pongo abelii} and \emph{Escherichia coli}.
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

@@

At the time of writing, more models of human history have been implemented than for
any other organism. These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth), the three population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations, the \cite{tennessen2012evolution} two population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth, the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components, a three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture, a complex model of ancient eurasian admixture from \cite{kamm2019efficiently},
and a synthetic model of oscillating population size from \cite{schiffels2014inferring}.
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{D. melanogaster}, three from \emph{A. thaliana}, and have under development
model for \emph{P. abelii} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample, as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{A. thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes. We have also implemented a two epoch and a three epoch model estimated from African
samples of \emph{A. thaliana} in \cite{huber2018gene}. 
%For \emph{P. pygmaeus} we have implemented the two
%species divergence model of \cite{locke2011comparative}, which models isolation with migration between \emph{P. pygmaeus}
%and \emph{P. abelii}. 
%For \emph{E. coli} we have a simple, large constant population size model implemented.
Beyond organism specific models \stdpopsim also includes a generic piecewise constant size model and
isolation with migration (IM) model which can be used with any genome and genetic map combination.

\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

>>>>

\subsection*{The Species Catalog}
The central feature of \stdpopsim is the species catalog, a systematic organization
of the key quantitative data needed to simulate a given species. These include a description
of the assembly, information about mutation rate, recombination rate(s), and generation time
in addition to a series of demographic models that are specific to that organism.

The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing, the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}, 
and we have models in development for \emph{Pongo abelii} and \emph{Escherichia coli}.
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

@@

At the time of writing, more models of human history have been implemented than for
any other organism (see Table \ref{tab:catalog}). 
These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth; Africa\_1T12), the three-population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations (OutOfAfrica\_3G09), the \cite{tennessen2012evolution} two-population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth (OutOfAfrica\_2T12), the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components (AmericanAdmixture\_4B11), a three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture (OutOfAfricaArchaicAdmixture\_5R19),
a complex model of ancient Eurasian admixture from \cite{kamm2019efficiently} (AncientEurasia\_9K19),
and a synthetic model of oscillating population size from \cite{schiffels2014inferring} (Zigzag\_1S14).
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{D. melanogaster}, three from \emph{A. thaliana}, and have under development
model for \emph{P. abelii} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample (African3Epoch\_1S16), as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations (OutOfAfrica\_2L06). For \emph{A. thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes (SouthMiddleAtlas\_1D17). We have also implemented a two-epoch and a three-epoch model estimated from African
samples of \emph{A. thaliana} in \cite{huber2018gene} (African2Epoch\_1H18 and African3Epoch\_1H18). 
%For \emph{P. pygmaeus} we have implemented the two
%species divergence model of \cite{locke2011comparative}, which models isolation with migration between \emph{P. pygmaeus}
%and \emph{P. abelii}. 
%For \emph{E. coli} we have a simple, large constant population size model implemented.
Beyond organism specific models \stdpopsim also includes a generic piecewise constant size model and
isolation with migration (IM) model which can be used with any genome and genetic map combination.

We note that we have put a good deal of effort into coming up with a standardized naming
conventions for species, genetic maps, and demographic models that will enable long term stability
of unique identifiers used throughout \stdpopsim. This is an essential feature to guarantee reproducibility.
\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.
